True versus apparent shapes of bow shocks and bow waves

### 2017-11-11 papers/Paper1/sec-conic.tex
Finish sec 5.2 on true shapes of cantoids and ancantoids

The type of quadric surface can be characterized by the
\textit{quadric parameter}:
\begin{equation}
  \label{eq:Tq}
  \Q \equiv \sigma \frac{b^2} {a^2} \ , 
\end{equation}
where \(\Q < 0\) corresponds to open surfaces (hyperboloids) and
\(\Q > 0\) corresponds to closed surfaces (oblate spheroids with
\(\Q > 1\) and prolate spheroids with \(\Q < 1\)).  Special cases are
the sphere (\(\Q = 1\)) and the paraboloid (\(\Q = 0\)).
Alternatively, one can define a \textit{quadric angle}:
\(\thetaQ = \sigma \tan^{-1} (b/a)\), which is marked in
Figure~\ref{fig:conics}.  In the case of hyperboloids, the asymptotic
opening angle of the wings (\S~\ref{sec:plan-alat-bow} and
Fig.~\ref{fig:characteristic-radii}) is
\(\theta_\infty = \pi + \thetaQ\) (note that \(\thetaQ < 0\) in this case), and
the minimum slope angle is \(\alpha_{\mathrm{min}} = \abs{\thetaQ}\), see
discussion following equation~\eqref{eq:thetapar}.

>>>>

The type of quadric surface can be characterized by the
\textit{quadric parameter}:
\begin{equation}
  \label{eq:Tq}
  \Q \equiv \sigma \frac{b^2} {a^2} \ , 
\end{equation}
where \(\Q < 0\) corresponds to open surfaces (hyperboloids) and
\(\Q > 0\) corresponds to closed surfaces (oblate spheroids with
\(\Q > 1\) and prolate spheroids with \(\Q < 1\)).  Special cases are
the sphere (\(\Q = 1\)) and the paraboloid (\(\Q = 0\)).
Alternatively, one can define a \textit{quadric angle}:
\begin{equation}
  \label{eq:thetaQ}
  \thetaQ = \sigma \tan^{-1} (b/a) \ ,
\end{equation}
which is marked in Figure~\ref{fig:conics}.  In the case of
hyperboloids, the asymptotic opening angle of the wings
(\S~\ref{sec:plan-alat-bow} and Fig.~\ref{fig:characteristic-radii})
is \(\theta_\infty = \pi + \thetaQ\) (note that \(\thetaQ < 0\) in this case), and
the minimum slope angle is \(\alpha_{\mathrm{min}} = \abs{\thetaQ}\), see
discussion following equation~\eqref{eq:thetapar}.

### 2017-11-11 papers/Paper1/sec-thin-shell.tex
Finish sec 5.2 on true shapes of cantoids and ancantoids

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/ancantoid-Pi-lambda-true}
  \caption{Shapes of cantoids and ancantoids in the
    \(\Pi\)--\(\Lambda\) plane.  Line colors correspond to different anisotropy
    indices, and match the values if Fig.~\ref{fig:r-beta}. Circle
    symbols mark particular \(\beta\) values: \(0, 0.01, 0.1\), from
    largest to smallest circle.  Square symbols mark \(\beta = 0.5\), but
    with \(\Lambda\) calculated exactly, instead of using the approximation
    of equation~\eqref{eq:Lambda-approx}.  The white plus symbol marks
    the result for the wilkinoid:
    \((\Pi, \Lambda) = (\frac53, \sqrt{3})\).  Background shading indicates
    the domains of different quadric classes: hyperboloids (white),
    prolate spheroids (darg gray), and oblate spheroids (light gray).}
  \label{fig:ancantoid-Pi-lambda-true}
\end{figure}

@@

The shapes of the ancantoid bow shocks are shown in
Figure~\ref{fig:r-beta} for three different values of \(\beta\), and are
compared with the \CRW{} results for cantoids (dashed curves).  The
location of these shapes in the planitude--alatude plane is shown in
Figure~\ref{fig:ancantoid-Pi-lambda-true}, where the gray background
shading indicates the zones of different quadric classes, as in
\S~\ref{sec:conic}, Figures~\ref{fig:quadric-projection-continued}
and~\ref{fig:projected-R90-Rc-snapshots}.  The \(k=0\) ancantoid is
identical to the cantoid for \(\theta \le 90^\circ\) (\(z > 0\)), but is slightly
more swept back in the far wings.\footnote{%
  Due to the discontinuity in the inner wind density at
  \(theta = 90^\circ\) (see Fig.~\ref{fig:anisotropic-arrows}), there is a
  discontinuity in the second derivative of the bow shape.} %
Since the true planitude and alatude depend on \(R(\theta)\) only in the
range \(\theta = [0, 90^\circ]\), the cantoid and the \(k = 0\) ancantoid
behave identically in Figure~\ref{fig:ancantoid-Pi-lambda-true}.

There is a general tendency for the bows to be flatter and more open
with increasing \(\beta\) and decreasing \(k\), with the cantoid being
most open at a given \(\beta\).  All the models cluster close to the
diagonal \(\Lambda \simeq \Pi\) in the planitude--alatude plane, but with a tendency
for \(\Lambda > \Pi\) at higher anisotropy.  There is therefore a degeneracy
between \(\beta\) and \(k\) for higher values of \(\beta\).  The wilkinoid
shape, which corresponds to the \(\beta \to 0\) limit of the cantoids is
marked by a white plus symbol in
Figure~\ref{fig:ancantoid-Pi-lambda-true}, which lies in the prolate
ellipsoid region of the plane.

@@

@@

>>>>

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/ancantoid-Pi-lambda-true}
  \caption{True shapes of cantoids and ancantoids in the
    \(\Pi\)--\(\Lambda\) plane, calculated according to results of
    App.~\ref{sec:thin-shell-shapes}.  For each line, \(\beta\) varies
    over the range \([0, 1]\) from lower left to upper right (although
    the black and red lines are truncated on the upper right), and
    line colors correspond to different anisotropy indices, matching
    those used in Fig.~\ref{fig:r-beta}. Circle symbols mark
    particular \(\beta\) values: \(0, 0.01, 0.1\), from largest to
    smallest circle.  Square symbols mark \(\beta = 0.5\), but with
    \(\Lambda\) calculated exactly, instead of using the approximation of
    equation~\eqref{eq:Lambda-approx}.  The white plus symbol marks
    the result for the wilkinoid:
    \((\Pi, \Lambda) = (\frac53, \sqrt{3})\).  Background shading indicates
    the domains of different quadric classes: hyperboloids (white),
    prolate spheroids (dark gray), and oblate spheroids (light gray).}
  \label{fig:ancantoid-Pi-lambda-true}
\end{figure}

@@

The shapes of the ancantoid bow shocks are shown in
Figure~\ref{fig:r-beta} for three different values of \(\beta\), and are
compared with the \CRW{} results for cantoids (dashed curves).  The
location of these shapes in the planitude--alatude plane is shown in
Figure~\ref{fig:ancantoid-Pi-lambda-true}, where the gray background
shading indicates the zones of different quadric classes, as in
\S~\ref{sec:conic}, Figures~\ref{fig:quadric-projection-continued}
and~\ref{fig:projected-R90-Rc-snapshots}.  Values of \(\Pi\) and
\(\Lambda\) are calculated via the analytic expressions derived in
Appendix~\ref{sec:ancantoid-planitude} and
\ref{sec:ancantoid-alatude}, respectively, which are only approximate
in the case of \(\Lambda\).  However, the filled square symbols show the
exact results for \(\beta = 0.5\), which can be seen to lie extremely close
to the approximate results, even for the worst case of \(k = 0\). The
leading term in the relative error of
equation~\eqref{eq:Lambda-approx} scales as \((\beta / (k + 2))^2\), so
the approximation is even better for smaller \(\beta\) and larger \(k\).

It is apparent from Figure~\ref{fig:r-beta} that the \(k=0\) ancantoid
is identical to the cantoid for \(\theta \le 90^\circ\) (\(z > 0\), to the right
of vertical dotted line in Fig.~\ref{fig:r-beta}), but is slightly
more swept back in the far wings.\footnote{%
  Due to the discontinuity in the inner wind density at
  \(\theta = 90^\circ\) (see Fig.~\ref{fig:anisotropic-arrows}), there is a
  discontinuity in the second derivative of the bow shape.} %
Since the true planitude and alatude depend on \(R(\theta)\) only in the
range \(\theta = [0, 90^\circ]\), the cantoid and the \(k = 0\) ancantoid
behave identically in Figure~\ref{fig:ancantoid-Pi-lambda-true}.
There is a general tendency for the bows to be flatter and more open
with increasing \(\beta\) and decreasing \(k\), with the cantoid being
most open at a given \(\beta\).  All the models cluster close to the
diagonal \(\Lambda \simeq \Pi\) in the planitude--alatude plane, but with a tendency
for \(\Lambda > \Pi\) at higher anisotropy.  There is therefore a degeneracy
between \(\beta\) and \(k\) for higher values of \(\beta\).  The wilkinoid
shape, which corresponds to the \(\beta \to 0\) limit of the cantoids, is
marked by a white plus symbol in
Figure~\ref{fig:ancantoid-Pi-lambda-true}, and lies in the prolate
spheroid region of the plane.  Cantoids lie either in the prolate
spheroid or hyperboloid regions, according to whether \(\beta\) is less
than or greater than about \(0.01\).  For ancantoids of increasing
\(k\), this dividing point moves to higher values of \(\beta\), until
almost the entire range of models with \(k = 8\) are within the
prolate spheroid zone.

@@

However, the true planitude and alatude, which are what would be
observed for a side-on viewing angle (\(i = 0\)), are not at all
sensitive to the behavior of the far wings of the bow shock, which has
a rather different implication as to which variety of quadric best
approximates each shape.  We illustrate this is
Figure~\ref{fig:ancantoid-angles}, which shows two different ways of
estimating the quadric angle, \(\theta_{\Q}\) (see \S~\ref{sec:conic}).
The first is from \((\Pi, \Lambda)\), as in
Figure~\ref{fig:ancantoid-Pi-lambda-true}:
\newcommand\head{^{\text{head}}}
\newcommand\tail{^{\text{tail}}}
\begin{equation}
  \label{eq:thetaQ-head}
  \theta_{\Q}\head =
  \sgn{\bigl(2 \Pi - \Lambda^2\bigr)} \;
  \tan^{-1} \Abs{2\Pi - \Lambda^2}^{1/2} \ ,
\end{equation}
which follows from equations~\eqref{eq:Tq}, \eqref{eq:thetaQ}, and
\eqref{eq:Tq-from-Pi-Lambda}.  The second is from the asymptotic
opening angle of the wings, \(\theta_\infty\) (Fig.~\ref{fig:characteristic-radii}):
\begin{equation}
  \label{eq:thetaQ-tail}
  \theta_{\Q}\tail = \theta_\infty - 180^\circ \ , 
\end{equation}
where \(\theta_\infty\) is calculated from
equation~\eqref{eq:anacantoid-theta-inf} for ancantoids, or
\eqref{eq:cantoid-theta-inf} for cantoids.  If the bow shock shape
were truly a quadric, then these two definitions would agree.
However, as seen in Figure~\ref{fig:ancantoid-angles}, this is not the
case for the cantoids and ancantoids.  While
\(\smash[b]{\theta_{\Q}\head}\) generally corresponds to a prolate spheroid
(except for the largest values of \(\beta\)),
\(\smash[b]{\theta_{\Q}\tail}\) always corresponds to a hyperbola.  This
has important implications for the projected shapes (as we will see in
the next section), since the far wings influence the projected
planitude and alatude when the inclination is large.

@@

\subsection{Apparent shapes of projected cantoids and ancantoids}
\label{sec:proj-shap-cant}
